Optimizing and Incrementalizing Collection Queries by AST Transformations

### 2017-04-15 chap-intro.tex
Add TODO on excluded papers (#15)

The second part makes the following contributions:
\begin{itemize}
\item We present a novel mathematical theory of changes and derivatives, which is more
  general than other work in the field because changes are
  first-class entities, they are distinct from base values and
  they are defined also for functions (\cref{sec:1st-order-changes}).
  %KO: I think the next sentence cannot be understood at this point.
  %We introduce changes for complex types, defined compositionally.
%
\item We present the first approach to incremental computation for pure
  $\lambda$-calculi by a source-to-source transformation, $\DERIVE$, that
  requires no run-time support. The transformation produces an incremental
  program in the same language; all optimization techniques for the original
  program are applicable to the incremental program as well.
%KO: commented this out. I think the purity is not important enough
%to deserve another sentence here, since we only vaguely hint
%at "further research".
%Since our incremental programs use no impure features, they are
%especially amenable to further optimizations, making this approach
%very suitable for further research.
%
% KO: Let's have one bullet point per section. Also, a conjecture
% sounds like a rather weak contribution
%\item We argue that incrementalization is efficient on
%  \emph{self-maintainable programs}, and discuss how further research on
%  static or dynamic memoization can speed up a larger class of programs (\cref{sec:performance-cons}).
%  \pg{This contribution references text which is now commented
%    out. I believe the text should be brought back in.}
%
We prove that our incrementalizing transformation $\DERIVE$
is correct~(\cref{eq:correctness})
by a machine-checked formalization in Agda~\citep{agda-head}.
The proof gives insight into the definition of $\DERIVE$: we
first construct the derivative $\EvalInc{-}$ of the denotational
semantics of a simply-typed $\lambda$-calculus term, that is, its
\emph{change semantics}.
%
Then, we show that $\DERIVE$ is produced by erasing
$\EvalInc{-}$ to a simply-typed program (\cref{sec:correctness}).

@@

\end{itemize}
% Modern libraries offer high-level APIs for manipulating (in-memory) collections;
% equivalently, such an API is an embedded domain-specific language (EDSL).

>>>>

\pg{rewrite this once we have a second part.}
The second part makes the following contributions:
\begin{itemize}
\item We present a novel mathematical theory of changes and derivatives, which is more
  general than other work in the field because changes are
  first-class entities, they are distinct from base values and
  they are defined also for functions.\pg{readd some reference.}
  %(\cref{sec:1st-order-changes}).
  %KO: I think the next sentence cannot be understood at this point.
  %We introduce changes for complex types, defined compositionally.
%
\item We present the first approach to incremental computation for pure
  $\lambda$-calculi by a source-to-source transformation, $\DERIVE$, that
  requires no run-time support. The transformation produces an incremental
  program in the same language; all optimization techniques for the original
  program are applicable to the incremental program as well.
%KO: commented this out. I think the purity is not important enough
%to deserve another sentence here, since we only vaguely hint
%at "further research".
%Since our incremental programs use no impure features, they are
%especially amenable to further optimizations, making this approach
%very suitable for further research.
%
% KO: Let's have one bullet point per section. Also, a conjecture
% sounds like a rather weak contribution
%\item We argue that incrementalization is efficient on
%  \emph{self-maintainable programs}, and discuss how further research on
%  static or dynamic memoization can speed up a larger class of programs (\cref{sec:performance-cons}).
%  \pg{This contribution references text which is now commented
%    out. I believe the text should be brought back in.}
%
We prove that our incrementalizing transformation $\DERIVE$
is correct~(\cref{eq:correctness})
by a machine-checked formalization in Agda~\citep{agda-head}.
The proof gives insight into the definition of $\DERIVE$: we
first construct the derivative $\EvalInc{-}$ of the denotational
semantics of a simply-typed $\lambda$-calculus term, that is, its
\emph{change semantics}.
%
Then, we show that $\DERIVE$ is produced by erasing
$\EvalInc{-}$ to a simply-typed program.% (\cref{sec:correctness}).

@@

\end{itemize}

\pg{Readdd section on excluded papers.
% \section{Excluded papers}
% During my PhD work I collaborated on several other papers.

% \paragraph{Software Product Lines}
% Initial work on Software Product Lines lead to \cite{KPO:VaMoS11} and
% \cite{KGREOB:OOPSLA11}. On the same line I collaborated to experiments in \cite{SRKGAK:SPLC11} and its updated version\pg{TODO}.
% \pg{Cite work on VarPL? Unlikely.}

% \paragraph{Modularity}
% \cite{OGKR:ECOOP11}

% \paragraph{DSL work}
% While working on modular embedded DSLs I collaborated to \citet{ErdwegGR12},
% \pg{TODO GADT paper}, \pg{TODO maybe manuscript on cake pattern?}.
}

% Modern libraries offer high-level APIs for manipulating (in-memory) collections;
% equivalently, such an API is an embedded domain-specific language (EDSL).
